---
title: "AirBNB analysis"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(sf)
library(tigris)
library(olsrr)
library(modelr)

library(osmdata)
library(spatstat)
library(ggspatial)
library(mapview)


```

### Loading in Data and Saving

```{r}
#state_boundaries <- states(year = 2025)
```

```{r}
# state_boundaries <- state_boundaries |> 
#  filter(REGION != 9 & !(STUSPS %in% c("AK", "HI")))
# us_coastline <- state_boundaries |> 
#   st_union() |> 
#   st_cast("MULTILINESTRING")

```

```{r}
# bnb_data <- read_csv("data/train.csv") |> 
#   select(-thumbnail_url)
# 
# bnb_sample <- bnb_data |> 
#   slice_sample(n = 10000)
```

```{r}
#save(state_boundaries, bnb_data, bnb_sample, us_coastline, file = "data/airbnb.RData")
```

### Exploratory Data Analysis

```{r}
load("data/airbnb.RData")
bnb_sample_sf <- st_as_sf(bnb_sample, coords = c("longitude", "latitude"), crs = 4326)
```

```{r}
bnb_sample_sf |> 
  ggplot() +
  geom_sf(data = state_boundaries) +
  geom_sf()
```

```{r}
us_coastline <- st_transform(us_coastline, 4326)
bnb_sample_sf <- bnb_sample_sf |> 
  mutate(price = exp(log_price),
         host_year = year(host_since)) |> 
  relocate(price, .before = log_price) |> 
  mutate(
    coast_distance = st_distance(geometry, us_coastline),
    coast_distance_km = (as.numeric(coast_distance)/1000)
    ) |> 
  mutate(
    clean_ammenities = str_replace_all(str_to_lower(amenities), "\"", "") |> 
      str_remove_all("\\{|\\}"), 
    ammenities_list = str_split(clean_ammenities, "\\s*,\\s*") |> 
      map(function(x) str_trim(x)),
    has_pool = map_lgl(ammenities_list, function(x) any(str_detect(x, "pools?")))
    ) |> 
  select(-clean_ammenities, -amenities)
```

## Models

### Initial Model with City only

```{r}
lm_bnb_0 <- lm(log(price) ~ city, data = bnb_sample_sf) 
summary(lm_bnb_0)
```

City already explains a large amount of variation in price

### Model with Room type and Host

```{r}
lm_bnb <- lm(log(price) ~ room_type + city + bedrooms + review_scores_rating, data = bnb_sample_sf) 
summary(lm_bnb)
```

R-squared moves up to 0.5644.

#### Adding an Interaction

```{r}
lm_bnb_2 <- lm(log(price) ~ room_type + city*log(coast_distance_km) + bedrooms + review_scores_rating, data = bnb_sample_sf) 
```

```{r}
summary(lm_bnb_2)
```

```{r}
plot(lm_bnb_2)
```

Chicago and DC have `coast_distance_km` that are too high. Including an interaction does not make sense here because the values of `coast_distance_km` do not overlap. Cities without coastal offerings will be removed from the analysis

#### Cities included in the final model

```{r}
bnb_sample_sf_coast <- bnb_sample_sf |> 
   filter(city %in% c("Boston", "LA", "SF", "NYC"))
```

```{r}
bnb_sample_sf |> 
  ggplot(aes(x = coast_distance_km, y = log(price), color = city)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_d() +
  geom_smooth(formula = y ~ x, method = "lm")
```

```{r}
bnb_sample_sf_coast |> 
  ggplot(aes(x = coast_distance_km, y = log(price), color = city)) +
  geom_point(alpha = 0.3) +
  scale_color_viridis_d() +
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  labs(
    x = "Distance to Coast (in km)",
    title = "Scatterplot of log(price) vs Distance to Coast by City",
    subtitle = "(for 8739 Listings)"
  ) +
  theme_gray()
```

```{r}
bnb_sample_sf_coast |> 
  ggplot(aes(x = coast_distance_km)) +
  geom_histogram(aes(fill = city), bins = 40) +
  scale_fill_viridis_d() +
  labs(
    title = "Histogram of Distance to Coast by City",
    x = "Distance to Coast (in km)",
    y = "Frequency",
    subtitle = "(for 8739 Listings)"
  ) + 
  theme_bw()
```

## Initial Model with Coastal Cities Only

```{r}
lm_bnb_coast <- lm(log(price) ~ room_type + city * log(coast_distance_km) + bedrooms + review_scores_rating, data = bnb_sample_sf_coast) 
```

```{r}
summary(lm_bnb_coast)
```

```{r}
plot(lm_bnb_coast)
```

```{r}
plot(~ log(price) + log(coast_distance_km) + bedrooms + review_scores_rating, data = bnb_sample_sf_coast)
```

```{r}
lm_bnb_coast_2 <- lm(log(price) ~ room_type + city * log(coast_distance_km) + bedrooms + review_scores_rating + has_pool, data = bnb_sample_sf_coast) 
```

```{r}
summary(lm_bnb_coast_2)
```
