---
title: "AirBNB analysis"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(sf)
library(tigris)
library(olsrr)
library(modelr)

library(osmdata)
library(spatstat)
library(ggspatial)
library(mapview)
chi_entertainment_data <- readRDS("data/chi_entertainment.rds")
chi_culture_data <- readRDS("data/chi_culture.rds")
chi_transport_data <- readRDS("data/chi_transport.rds")
chi_park_data <- readRDS("data/chi_parks.rds")


```

### Loading in Data and Saving

```{r}
#state_boundaries <- states(year = 2025)
```

```{r}
# state_boundaries <- state_boundaries |> 
#  filter(REGION != 9 & !(STUSPS %in% c("AK", "HI")))
# us_coastline <- state_boundaries |> 
#   st_union() |> 
#   st_cast("MULTILINESTRING")

```

```{r}
# bnb_data <- read_csv("data/train.csv") |> 
#   select(-thumbnail_url)
# 
# bnb_sample <- bnb_data |> 
#   slice_sample(n = 10000)
```

```{r}
#save(state_boundaries, bnb_data, bnb_sample, us_coastline, file = "data/airbnb.RData")
```

### Exploratory Data Analysis

```{r}
load("data_private/airbnb.RData")
bnb_sample_sf <- st_as_sf(bnb_sample, coords = c("longitude", "latitude"), crs = 4326)
```

```{r}
bnb_sample_sf |> 
  ggplot() +
  geom_sf(data = state_boundaries) +
  geom_sf()
```

```{r}
us_coastline <- st_transform(us_coastline, 4326)
bnb_sample_sf <- bnb_sample_sf |> 
  mutate(price = exp(log_price),
         host_year = year(host_since)) |> 
  relocate(price, .before = log_price) |> 
  mutate(
    coast_distance = st_distance(geometry, us_coastline),
    coast_distance_km = (as.numeric(coast_distance)/1000)
    ) |> 
  mutate(
    clean_ammenities = str_replace_all(str_to_lower(amenities), "\"", "") |> 
      str_remove_all("\\{|\\}"), 
    ammenities_list = str_split(clean_ammenities, "\\s*,\\s*") |> 
      map(function(x) str_trim(x)),
    has_pool = map_lgl(ammenities_list, function(x) any(str_detect(x, "pools?")))
    ) |> 
  select(-clean_ammenities, -amenities)
```

## Models

### Initial Model with City only

```{r}
lm_bnb_0 <- lm(log(price) ~ city, data = bnb_sample_sf) 
summary(lm_bnb_0)
```

City already explains a large amount of variation in price

### Model with Room type and Host

```{r}
lm_bnb <- lm(log(price) ~ room_type + city + bedrooms + review_scores_rating, data = bnb_sample_sf) 
summary(lm_bnb)
```

R-squared moves up to 0.5644.

#### Adding an Interaction

```{r}
lm_bnb_2 <- lm(log(price) ~ room_type + city*log(coast_distance_km) + bedrooms + review_scores_rating, data = bnb_sample_sf) 
```

```{r}
summary(lm_bnb_2)
```

```{r}
plot(lm_bnb_2)
```

Chicago and DC have `coast_distance_km` that are too high. Including an interaction does not make sense here because the values of `coast_distance_km` do not overlap. Cities without coastal offerings will be removed from the analysis

#### Cities included in the final model

```{r}
bnb_sample_sf_coast <- bnb_sample_sf |> 
   filter(city %in% c("Boston", "LA", "SF", "NYC"))
```

```{r}
bnb_sample_sf |> 
  ggplot(aes(x = coast_distance_km, y = log(price), color = city)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_d() +
  geom_smooth(formula = y ~ x, method = "lm")
```

```{r}
bnb_sample_sf_coast |> 
  ggplot(aes(x = coast_distance_km, y = log(price), color = city)) +
  geom_point(alpha = 0.3) +
  scale_color_viridis_d() +
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  labs(
    x = "Distance to Coast (in km)",
    title = "Scatterplot of log(price) vs Distance to Coast by City",
    subtitle = "(for 8739 Listings)"
  ) +
  theme_gray()
```

```{r}
bnb_sample_sf_coast |> 
  ggplot(aes(x = coast_distance_km)) +
  geom_histogram(aes(fill = city), bins = 40) +
  scale_fill_viridis_d() +
  labs(
    title = "Histogram of Distance to Coast by City",
    x = "Distance to Coast (in km)",
    y = "Frequency",
    subtitle = "(for 8739 Listings)"
  ) + 
  theme_bw()
```

## Initial Model with Coastal Cities Only

```{r}
lm_bnb_coast <- lm(log(price) ~ room_type + city * log(coast_distance_km) + bedrooms + review_scores_rating, data = bnb_sample_sf_coast) 
```

```{r}
summary(lm_bnb_coast)
```

```{r}
plot(lm_bnb_coast)
```

```{r}
plot(~ log(price) + log(coast_distance_km) + bedrooms + review_scores_rating, data = bnb_sample_sf_coast)
```

```{r}
lm_bnb_coast_2 <- lm(log(price) ~ room_type + city * log(coast_distance_km) + bedrooms + review_scores_rating + has_pool, data = bnb_sample_sf_coast) 
```

```{r}
summary(lm_bnb_coast_2)
```

### OSM Data

```{r}
chi_bnb <- bnb_sample_sf |> 
  filter(city == "Chicago")
chi_bbox <-st_bbox(chi_bnb) |> 
  st_as_sfc()
chi_bbox_buffered <- chi_bbox |> 
  st_transform(26916) |>  #code for projection to meters in chicago 
  st_buffer(dist = 5000) |> 
  st_transform(4326)


```

```{r}
ggplot() +
  annotation_map_tile() + 
  geom_sf(data = chi_bbox_buffered, alpha = 0.3) +
  geom_sf(data = chi_bbox, alpha = 0.3) +
  geom_sf(data = chi_bnb)
```

```{r}
entertainment_types <- c("bar",
                         "pub",
                         "nightclub",
                         "theatre",
                         "cinema",
                         "arts_centre"
                         )

culture_types <- c("museum",
                  "gallery",
                 "attraction",
                  "hotel"
                 )

transport_types <- c("station", 
                     "stop_position"
                     )
```

#### Code to get data from OSM and save RDS files
```{r}
# entertainment_query <- opq(bbox = chi_bbox_buffered) |>
#   add_osm_feature(
#     key = "amenity",
#     value = c(
#       "bar",
#       "pub",
#       "nightclub",
#       "theatre",
#       "cinema",
#       "arts_centre"
#     )
#   )
# 
# entertainment_data <- osmdata_sf(entertainment_query)
```

```{r}
# culture_query <- opq(bbox = chi_bbox_buffered) |>
#   add_osm_feature(
#     key = "tourism",
#     value = c(
#       "museum",
#       "gallery",
#       "attraction",
#       "viewpoint",
#       "hotel"
#     )
#   )
# 
# culture_data <- osmdata_sf(culture_query)
```

```{r}
# transport_query <- opq(bbox = chi_bbox_buffered) |>
#   add_osm_feature(
#     key = "public_transport",
#     value = c("station", "stop_position")
#   )
# 
# transport_data <- osmdata_sf(transport_query)
```

```{r}
# saveRDS(entertainment_data, file = "data/chi_entertainment.rds")
# saveRDS(culture_data, file = "data/chi_culture.rds")
# saveRDS(transport_data, file = "data/chi_transport.rds")

```

#### Functions to clean OSM Data

Note: There is currently a warning message for this function because `st_point_on_surface()` should get spatial data that is not in WGS84
```{r}
clean_osm_data <- function(osm_layer, tag_column, tags, source_type = c("point", "polygon", "multipolygon", "lines")){
  if(is.null(osm_layer)){
    return(NULL)
  }
  
  source_type <- match.arg(source_type)
  
  cleaned_data <- osm_layer |> 
      filter(.data[[tag_column]] %in% tags) |> 
      st_point_on_surface() |> 
      mutate(source = source_type) |> 
      select(
      any_of(c("osm_id", "name", "geometry", "source", tag_column)),
      starts_with("addr:")
    )
  return(cleaned_data)
}


aggregate_osm_data <- function(osm_dataset, tag_column, tags){
  
  venue_points <- clean_osm_data(osm_dataset$osm_points, tag_column = tag_column, tags = tags, source_type = "point")
  venue_poly <- clean_osm_data(osm_dataset$osm_polygons, tag_column = tag_column, tags = tags, source_type = "polygon")
  venue_multipoly <- clean_osm_data(osm_dataset$osm_multipolygons, tag_column = tag_column, tags = tags, source_type = "multipolygon")
  venue_lines <- clean_osm_data(osm_dataset$osm_lines, tag_column = tag_column, tags = tags, source_type = "lines")
  city_venues <- bind_rows(
    venue_points,
    venue_poly,
    venue_multipoly,
    venue_lines
  )
  return(city_venues)
 
}
```
```{r}
chi_entertainment_clean <- aggregate_osm_data(chi_entertainment_data, "amenity", entertainment_types) 
chi_culture_clean <- aggregate_osm_data(chi_culture_data, "tourism", culture_types) 
chi_transport_clean <- aggregate_osm_data(chi_transport_data, "public_transport", transport_types) 
```

```{r}
saveRDS(chi_entertainment_clean, "data/chi_entertainment_clean.rds")
saveRDS(chi_culture_clean, "data/chi_culture_clean.rds")
saveRDS(chi_transport_clean, "data/chi_transport_clean.rds")
```














